
---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-")
```

#lexvarsdatr

A collection of psycholinguistic/behavioral data, collated from supplemental materials and public databases. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(lexvarsdatr) #devtools::install_github("jaytimm/lexvarsdatr")
```

Data included in package:
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
Data <- c("CELEX - English lemma database","Lexical decision and naming","Concreteness ratings","AoA ratings", "Word association")


Source <- c("Baayen, R. H., Piepenbrock, R., & Gulikers, L. (1995). The CELEX lexical database [webcelex]. *Philadelphia, PA: University of Pennsylvania, Linguistic Data Consortium*.", "Balota, D. A., Yap, M. J., Hutchison, K. A., Cortese, M. J., Kessler, B., Loftis, B., ... & Treiman, R. (2007). The English lexicon project. *Behavior research methods*, 39(3), 445-459.","Brysbaert, M., Warriner, A. B., & Kuperman, V. (2014). Concreteness ratings for 40 thousand generally known English word lemmas. *Behavior research methods*, 46(3), 904-911.","Kuperman, V., Stadthagen-Gonzalez, H., & Brysbaert, M. (2012). Age-of-acquisition ratings for 30,000 English words. *Behavior Research Methods*, 44(4), 978-990.","Nelson, D. L., McEvoy, C. L., & Schreiber, T. A. (2004). The University of South Florida free association, rhyme, and word fragment norms. *Behavior Research Methods, Instruments, & Computers*, 36(3), 402-407.")

kable(data.frame(Data=Data,Source=Source))
```



###lvdr_behav_data

For convenience, response times in lexical decision/naming, concreteness ratings, and AoA ratings have been collated into a single data frame, `lex_behav_data`.  Approximately 18K word forms are included in all three data sets.  

```{r message=FALSE, warning=FALSE}
library(tidyverse)
lexvarsdatr::lvdr_behav_data %>%
  na.omit %>%
  head
```



###lvdr_celex

The English lemma portion of the CELEX database can be accessed as `lvdr_celex`.  Morphological families can be extracted for both word forms and affixes with the `lvdr_get_family` function.  
```{r}
lexvarsdatr::lvdr_get_family(form="think",multiword = TRUE)
```

```{r}
lexvarsdatr::lvdr_get_family(type="SUF",form="wise")
```


###lvdr_association

The South Florida word association data set lives in `lvdr_association`.  A description of variables included in the normed data set, as well as methodologies, can be found [here](http://w3.usf.edu/FreeAssociation/).  The `lvdr_get_associates` function enables quick access to word associates for a given cue; associates are listed in descending order (per total subject responses).

```{r}
lexvarsdatr::lvdr_get_associates(cue='think')
```

Additional querying functionality for both the CELEX and word association data sets is forthcoming.
