
---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "README-")
```

##lexvarsdatr: 

Baayen, R. H., Piepenbrock, R., & Gulikers, L. (1995). The CELEX lexical database [webcelex]. *Philadelphia, PA: University of Pennsylvania, Linguistic Data Consortium*.

Balota, D. A., Yap, M. J., Hutchison, K. A., Cortese, M. J., Kessler, B., Loftis, B., ... & Treiman, R. (2007). The English lexicon project. *Behavior research methods*, 39(3), 445-459.

Brysbaert, M., Warriner, A. B., & Kuperman, V. (2014). Concreteness ratings for 40 thousand generally known English word lemmas. *Behavior research methods*, 46(3), 904-911.

Kuperman, V., Stadthagen-Gonzalez, H., & Brysbaert, M. (2012). Age-of-acquisition ratings for 30,000 English words. *Behavior Research Methods*, 44(4), 978-990.

Nelson, D. L., McEvoy, C. L., & Schreiber, T. A. (2004). The University of South Florida free association, rhyme, and word fragment norms. *Behavior Research Methods, Instruments, & Computers*, 36(3), 402-407.




```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
#devtools::install_github("jaytimm/corpuslingr")
library(corpuslingr)
```

